Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-24 arxiv/diff_cov.tex
wip

We first provide a key technical tool that we derived using recent developments in random matrix theory \cite{??}.
To motivate our study, we consider a specialized setting where the output layer weights $A_1, A_2$ are both equal to one.
Without loss of generality, we consider the prediction loss of hard parameter sharing for task two and the same analysis applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align*}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align*}
For the prediction loss, since the empirical loss is close to the expected loss similar to what we have shown in Claim \ref{claim_pred_err}, we consider the expected loss conditional on the covariates as follows:
\begin{align*}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	\bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2
	+ \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}.
\end{align*}

Based on the above intuition, in the following we develop generalization bounds that relate the bias and variance of multi-task learning to the sample sizes and the covariance matrices of both tasks.
We introduce a key matrix $M\equiv M(W_1/W_2): =\frac{W_1}{W_2} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ that captures the covariate shift between task $1$ and task $2$.

\begin{definition}[Limiting variance]
	Define the following
	\[ \Pi_{\vari} \define \frac{1}{n_1 + n_2}\cdot \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2}. \]

	In the above equation, $a_1$ and $a_2$ are the solutions of the following two equations that only depend on the sample sizes $\rho_1, \rho_2$, and the singular values of the covariate shift matrix $M$:
	\begin{align}
		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
	\end{align}
\end{definition}
%We now state our main result for two tasks with both covariate and model shift. It shows that the information transfer is determined by two deterministic quantities $\Delta_{\bias}$ and $\Delta_{\vari}$, which give the change of model shift bias and the change of variance, respectively. The exact forms of $\Delta_{\bias}$ and $\Delta_{\vari}$ will be given in Lemma \ref{thm_model_shift}.

\begin{definition}[Limiting bias]
Recall that $a_1$ and $a_2$ are the solutions of the equation system \eqref{eq_a12extra}. Let
	\be\label{def Pibias} \Pi_\bias \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \ee
where $a_{3}$ and $a_4$ are the solutions of the following two equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
		= b_0 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
\end{definition}

\begin{theorem}[Two-task case]\label{thm_main_informal}
	%For the setting of two tasks, let $\delta > 0$ be a fixed error margin, $\rho_2 > 1$ and $\rho_1 \gtrsim \delta^{-2}\cdot \lambda_{\min}(M)^{-4} \norm{\Sigma_1} \max(\norm{\beta_1}^2, \norm{\beta_2}^2)$.
	For the setting of two tasks, let $C$ be a fixed constant.
	Let $B, W_1, W_2$ be any local minimum of equation \eqref{eq_mtl}.
 	%There exist two deterministic functions $\Delta_{\bias}$ and $\Delta_{\vari}$ that only depend on scaled model distance $\beta_1 - \frac{W_1}{W_2} \beta_2$, sample sizes $n_1 = \rho_1 \cdot p, n_2 = \rho_2 \cdot p$, and covariate shift matrix $M$ such that
	We have that
	\begin{align*}
		\bigabs{L(\hat{\beta}_2^{\MTL}) - \Pi_{\vari} - (\beta_1 - \frac{W_1}{W_2}\beta_2)^{\top}\Pi_{\bias} (\beta_1 - \frac{W_1}{W_2}\beta_2)} \le \frac{ C\cdot \max(\norm{\beta_1}, \norm{\beta_2}) \cdot \sqrt{\norm{\Sigma_1}} } {\lambda_{\min}^2(M)) } \cdot \frac{1}{\sqrt{\rho_1}}
	\end{align*}
%	\begin{enumerate}
%		\item[a)] \textbf{Positive transfer:} If $\Delta_{\bias} < \Delta_{\vari} -  \frac{\delta}{\sqrt{\rho_1}} $, then w.h.p. over the randomness of $X_1, X_2, \varepsilon_1, \varepsilon_2$, we have
%			\[ \te(\hat{\beta}_2^{\MTL}) < \te(\hat{\beta}_2^{\STL}).  \]
%		\item[b)] \textbf{Negative transfer:} If $\Delta_{\bias} > \Delta_{\vari} + \frac{\delta}{\sqrt{\rho_1}}$, then w.h.p. over the randomness of $X_1, X_2, \varepsilon_1, \varepsilon_2$, we have
%			\[ \te(\hat{\beta}_2^{\MTL}) > \te(\hat{\beta}_2^{\STL}). \]
%	\end{enumerate}
\end{theorem}
In words, the above result shows that a deterministic function $\Delta_{\bias} - \Delta_{\vari}$ determines whether the prediction loss of the empirical multi-task learning estimator is lower than that of single-task learning, up to an error that scales as $\delta / \sqrt{\rho_1}$.
We make several remarks about Theorem \ref{thm_main_informal}.
First, as the amount of data from the source task increases, we get more accurate predictions  since the error scales down.
This applies to many practical settings where collecting labeled data for the target task is expensive and auxillary (source) task data is easier to obtain.
%Theorem \ref{thm_main_informal} applies to settings where large amounts of source task data are available but the target sample size is small.
%For such settings, we obtain a sharp transition from positive transfer to negative transfer determined by $\Delta_{\bias} - \Delta_{\vari}$.
%determined by the covariate shift matrix and the model shift.
%The bounds get tighter and tighter as $\rho_1$ increases.
Second, the deterministic function $\Delta_{\bias} - \Delta_{\vari}$ depends on the three task properties that we care about and the precise form can be found in Section \ref{sec_proof_general}.
Finally, later on in Section \ref{sec_special}, we will study how varying each task property affects the performance of multi-task learning in depth.
%While the general form of these functions can be complex (as are previous generalization bounds for MTL), they admit interpretable forms for simplified settings.

@@

\paragraph{Key ingredients.}
We introduce two key lemmas that show tight asymptotic convergence rate of the bias equation \eqref{eq_bias_2task} and the variance equation  as sample sizes increase to infinity.
%The following two lemmas  are the main random matrix theoretical results of this paper, which will be used to estimate the two terms on the right-hand side of \eqref{eq_te_mtl_2task}.
For the covariate shift matrix $M$, let $\lambda_1, \lambda_2, \dots, \lambda_p$ be its singular values in descending order.
%which deals with the inverse of the sum of two random matrices, which
%any is can be viewed as a special case of Theorem \ref{thm_model_shift}.

\begin{lemma}[Variance asymptotics]\label{lem_cov_shift}
	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	%Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
	%Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and
	Let $\Sigma \in \real^{p \times p}$ be any fixed and deterministic matrix.
	In the setting of Theorem \ref{thm_main_informal},
	with probability $1-\oo(1)$ over the randomness of $X_1$ and $X_2$, we have that %for any constant $\e>0$,
	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma} = \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma} + o(\norm{\Sigma}), %\bigo{\|\| p^{-1/2+\epsilon}}
	\end{align}
where $(a_1,a_2)$ is the solution to equation \eqref{eq_a12extra} with $\lambda_i$ being the singular values of $M\equiv M(1): = \Sigma_1^{1/2}\Sigma_2^{-1/2}$.
\end{lemma}

Lemma \ref{lem_cov_shift} derives the asymptotic limit of the variance equation \eqref{eq_var_2task}.
To see this, we rescale $X_1$ with $W_1 / W_2$ and set the matrix $\Sigma$ as $\sigma^2 \Sigma_2$.
%allows us to get a tight bound on equation \eqref{eq_te_var}, that only depends on \textit{sample size}, \textit{covariate shift} and the scalar $\hat{v}$.
As a remark, the concentration error $o(\|A\|)$ on the right hand side of equation \eqref{lem_cov_shift_eq} reduces provided with stronger moment assumptions on $X_1$ and $X_2$. % p^{-1/2+\epsilon}
Recall from Section \ref{sec_prelim} that the feature vectors are generated by $\Sigma_i^{1/2} z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance.
Provided that for every entry of $z$, its $k$-th moment exists, then we can obtain a concentration error at most $\bigo{\norm{A} p^{-1/2 + 2/k + o(1)}}$.
\todo{We remark that one can probably derive the same asymptotic result using free probability theory (see e.g. \cite{nica2006lectures}), but our results \eqref{lem_cov_shift_eq} and \eqref{lem_cov_derv_eq} also give an almost sharp error bound $\bigo{ p^{-1/2+\epsilon}}$.}

Lemma \ref{lem_cov_shift} also implies the asymptotic limit of the prediction loss of single-task learning.
%The next lemma, which is , helps to determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})=\sigma^2   \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}$ as $p\to \infty$.
In particular, by setting $X_1 = 0$, we obtain Lemma \ref{lem_minv}.
%\begin{corollary}[See e.g. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
%	In the setting of Lemma \ref{lem_cov_shift}, with high probability we have that
%	\be\label{XXA}  \bigtr{(X_2^{\top}X_2)^{-1}\Sigma} = \frac{1}{n_2 - p} \cdot \bigtr{\Sigma_2^{-1}\Sigma} + o(\norm{\Sigma}). \ee
%\end{corollary} %\bigo{ \|A\|p^{-1/2+\epsilon}}
To see this, note that when $n_1=0$, $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ is the solution of \eqref{eq_a12extra}, and one can see that equation \eqref{lem_cov_shift_eq} reduces to equation \eqref{XXA}.
We briefly describe the history behind the above well-known result.
When the entries of $X_2$ are multivariate Gaussian, this result recovers the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For general non-Gaussian random matrices, it can be obtained using Stieltjes transform method; see e.g., Lemma 3.11 of \cite{bai2009spectral}.
Here we have stated a result from Theorem 2.4 in \cite{isotropic}, which gives an almost sharp concentration error bound.
One can see that our result extends Lemma \ref{lem_minv} from a single sample covariance matrix to the sum of two independent sample covariance matrices.

>>>>

We first provide a key technical tool that we derived using recent developments in random matrix theory \cite{??}.
To motivate our study, we consider a specialized setting where the output layer weights $A_1, A_2$ are both equal to one.
Without loss of generality, we consider the prediction loss of hard parameter sharing for task two and the same analysis applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align*}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align*}
For the prediction loss, the empirical loss is close to the expected loss similar to what we have shown in Claim \ref{claim_pred_err}.
Hence, we consider the expected loss conditional on the covariates as follows:
\begin{align*}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	\bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2
	+ \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}.
\end{align*}

Interestingly, the prediction loss admits a bias-variance decomposition similar to the same covariates setting.
In both the bias and variance equations, the matrix inverse $\hat{\Sigma}^{-1}$ is a crucial quantity.
Intuitively, this matrix inverse combines the sample covariance matrix of both tasks, and it is not hard to see that the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by the sample sizes.
However, since the two tasks have different covariates, the scaling of $\hat{\Sigma}^{-1}$ not only depends on the tasks' sample sizes, but also depends on the covariate shift between the two tasks.
We introduce a key result that show tight asymptotic convergence rate of the bias equation \eqref{eq_bias_2task} and the variance equation  as sample sizes increase to infinity.
%The following two lemmas  are the main random matrix theoretical results of this paper, which will be used to estimate the two terms on the right-hand side of \eqref{eq_te_mtl_2task}.
We define a key quantity $M \define \Sigma_1^{1/2}\Sigma_2^{-1/2}$ that captures the covariate shift between task $1$ and task $2$.
Let $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M$ in descending order.
%Based on the above intuition, in the following we develop generalization bounds that relate the bias and variance of multi-task learning to the sample sizes and the covariance matrices of both tasks.

\begin{theorem}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	Then we have the following asymptotic limits:
	\begin{enumerate}
		\item[i)] Limiting variance:
		 	Define the following
			\[ \Pi_{\vari} \define \frac{1}{n_1 + n_2}\cdot \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2}. \]
			In the above equation, $a_1$ and $a_2$ are the solutions of the following two equations that only depend on the sample sizes $\rho_1, \rho_2$, and the singular values of the covariate shift matrix $M$:
			\begin{align}
				a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
			\end{align}
			Then we have that
			\begin{align}\label{lem_cov_shift_eq}
				\bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} = \bigtr{\Pi_{\vari}} + \OO(p^{-c}\norm{\Sigma^{(2)}}), %\bigo{\|\| p^{-1/2+\epsilon}}
			\end{align}
		\item[ii)] Limiting bias: Let
			\be\label{def Pibias} \Pi_\bias \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \ee
				where $a_{3}$ and $a_4$ are the solutions of the following two equations:
			\begin{gather}\label{eq_a34extra}
				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
				= b_0 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
			\end{gather}
				In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
			Then, we have that
			\begin{align}\label{lem_cov_derv_eq}
				\bignorm{\Sigma_2^{1/2} \bigbrace{\frac{X_1^{\top}X_1 + X_2^{\top}X_2}{n_1 + n_2}}^{-1} \Sigma_1^{1/2} w}^2
						= w^{\top} \Pi_\bias w + o(\|w\|^2),
			\end{align}
	\end{enumerate}
\end{theorem}

Lemma \ref{lem_cov_shift} also implies the asymptotic limit of the prediction loss of single-task learning.
%The next lemma, which is , helps to determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})=\sigma^2   \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}$ as $p\to \infty$.
In particular, by setting $X_1 = 0$, we obtain Lemma \ref{lem_minv}.
%\begin{corollary}[See e.g. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
%	In the setting of Lemma \ref{lem_cov_shift}, with high probability we have that
%	\be\label{XXA}  \bigtr{(X_2^{\top}X_2)^{-1}\Sigma} = \frac{1}{n_2 - p} \cdot \bigtr{\Sigma_2^{-1}\Sigma} + o(\norm{\Sigma}). \ee
%\end{corollary} %\bigo{ \|A\|p^{-1/2+\epsilon}}
To see this, note that when $n_1=0$, $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ is the solution of \eqref{eq_a12extra}, and one can see that equation \eqref{lem_cov_shift_eq} reduces to equation \eqref{XXA}.
We briefly describe the history behind the above well-known result.
When the entries of $X_2$ are multivariate Gaussian, this result recovers the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For general non-Gaussian random matrices, it can be obtained using Stieltjes transform method; see e.g., Lemma 3.11 of \cite{bai2009spectral}.
Here we have stated a result from Theorem 2.4 in \cite{isotropic}, which gives an almost sharp concentration error bound.
One can see that our result extends Lemma \ref{lem_minv} from a single sample covariance matrix to the sum of two independent sample covariance matrices.
%We now state our main result for two tasks with both covariate and model shift. It shows that the information transfer is determined by two deterministic quantities $\Delta_{\bias}$ and $\Delta_{\vari}$, which give the change of model shift bias and the change of variance, respectively. The exact forms of $\Delta_{\bias}$ and $\Delta_{\vari}$ will be given in Lemma \ref{thm_model_shift}.

Interestingly, the asymptotic limit of the variance only depends on the sample sizes and the covariate shift matrix.
Next, we derive the asymptotic limit of the bias equation \eqref{eq_bias_2task}, which depends on the scaled model distance in addition to sample size and covariate shift. %(cf. Lemma \ref{lem_cov_derivative} in Appendix \ref{app_proof_main_thm}).

\begin{corollary}[Two tasks with different covariates]\label{thm_main_informal}
	%For the setting of two tasks, let $\delta > 0$ be a fixed error margin, $\rho_2 > 1$ and $\rho_1 \gtrsim \delta^{-2}\cdot \lambda_{\min}(M)^{-4} \norm{\Sigma_1} \max(\norm{\beta_1}^2, \norm{\beta_2}^2)$.
	For the setting of two tasks, let $C$ be a fixed constant.
	Let $B, W_1, W_2$ be any local minimum of equation \eqref{eq_mtl}.
 	%There exist two deterministic functions $\Delta_{\bias}$ and $\Delta_{\vari}$ that only depend on scaled model distance $\beta_1 - \frac{W_1}{W_2} \beta_2$, sample sizes $n_1 = \rho_1 \cdot p, n_2 = \rho_2 \cdot p$, and covariate shift matrix $M$ such that
	We have that
	\begin{align*}
		\bigabs{L(\hat{\beta}_2^{\MTL}) - \Pi_{\vari} - (\beta_1 - \frac{W_1}{W_2}\beta_2)^{\top}\Pi_{\bias} (\beta_1 - \frac{W_1}{W_2}\beta_2)} \le \frac{ C\cdot \max(\norm{\beta_1}, \norm{\beta_2}) \cdot \sqrt{\norm{\Sigma_1}} } {\lambda_{\min}^2(M)) } \cdot \frac{1}{\sqrt{\rho_1}}
	\end{align*}
%	\begin{enumerate}
%		\item[a)] \textbf{Positive transfer:} If $\Delta_{\bias} < \Delta_{\vari} -  \frac{\delta}{\sqrt{\rho_1}} $, then w.h.p. over the randomness of $X_1, X_2, \varepsilon_1, \varepsilon_2$, we have
%			\[ \te(\hat{\beta}_2^{\MTL}) < \te(\hat{\beta}_2^{\STL}).  \]
%		\item[b)] \textbf{Negative transfer:} If $\Delta_{\bias} > \Delta_{\vari} + \frac{\delta}{\sqrt{\rho_1}}$, then w.h.p. over the randomness of $X_1, X_2, \varepsilon_1, \varepsilon_2$, we have
%			\[ \te(\hat{\beta}_2^{\MTL}) > \te(\hat{\beta}_2^{\STL}). \]
%	\end{enumerate}
\end{corollary}
In words, the above result shows that a deterministic function $\Delta_{\bias} - \Delta_{\vari}$ determines whether the prediction loss of the empirical multi-task learning estimator is lower than that of single-task learning, up to an error that scales as $\delta / \sqrt{\rho_1}$.
We make several remarks about Corollary \ref{thm_main_informal}.
First, as the amount of data from the source task increases, we get more accurate predictions  since the error scales down.
This applies to many practical settings where collecting labeled data for the target task is expensive and auxillary (source) task data is easier to obtain.
%Theorem \ref{thm_main_informal} applies to settings where large amounts of source task data are available but the target sample size is small.
%For such settings, we obtain a sharp transition from positive transfer to negative transfer determined by $\Delta_{\bias} - \Delta_{\vari}$.
%determined by the covariate shift matrix and the model shift.
%The bounds get tighter and tighter as $\rho_1$ increases.
Second, the deterministic function $\Delta_{\bias} - \Delta_{\vari}$ depends on the three task properties that we care about and the precise form can be found in Section \ref{sec_proof_general}.
Finally, later on in Section \ref{sec_special}, we will study how varying each task property affects the performance of multi-task learning in depth.

Lemma \ref{lem_cov_shift} derives the asymptotic limit of the variance equation \eqref{eq_var_2task}.
To see this, we rescale $X_1$ with $W_1 / W_2$ and set the matrix $\Sigma$ as $\sigma^2 \Sigma_2$.
%allows us to get a tight bound on equation \eqref{eq_te_var}, that only depends on \textit{sample size}, \textit{covariate shift} and the scalar $\hat{v}$.
As a remark, the concentration error $o(\|A\|)$ on the right hand side of equation \eqref{lem_cov_shift_eq} reduces provided with stronger moment assumptions on $X_1$ and $X_2$. % p^{-1/2+\epsilon}
Recall from Section \ref{sec_prelim} that the feature vectors are generated by $\Sigma_i^{1/2} z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance.
Provided that for every entry of $z$, its $k$-th moment exists, then we can obtain a concentration error at most $\bigo{\norm{A} p^{-1/2 + 2/k + o(1)}}$.
\todo{We remark that one can probably derive the same asymptotic result using free probability theory (see e.g. \cite{nica2006lectures}), but our results \eqref{lem_cov_shift_eq} and \eqref{lem_cov_derv_eq} also give an almost sharp error bound $\bigo{ p^{-1/2+\epsilon}}$.}
%We will use the matrix fractional notation $\frac{X}{Y}$ for two PSD matrices that share the same eigenspace.
%That is, suppose that the SVD of $X, Y$ is $X = U D_1 U^{\top}$ and $Y = U D_2 U^{\top}$.
%We denote $\frac{X}{Y} = U D_{1} D_2^{-1} U^{\top}$.
%Using the notation, we introduce an important quantity that describes the limit of the bias equation as $p$ goes to infinity.
%Our next result shows that the bias equation \eqref{eq_bias_2task} is described by the above limiting bias asymptotically.
%\begin{lemma}[Bias asymptotics]\label{lem_cov_derivative}
%In the setting of Theorem \ref{thm_main_informal}, let $w \in \R^p$ be any vector that is independent of $X_1$ and $X_2$.
%With high probability, we have that
%\begin{equation}\label{lem_cov_derv_eq}
%\begin{split}
%\bignorm{\Sigma_2^{1/2} \bigbrace{\frac{X_1^{\top}X_1 + X_2^{\top}X_2}{n_1 + n_2}}^{-1} \Sigma_1^{1/2} w}^2
%= w^{\top} \Pi_\bias w + o(\|w\|^2),
%\end{split}
%\end{equation}
%where $\Pi_\bias$ was defined in \eqref{def Pibias} with $M\equiv M(1): = \Sigma_1^{1/2}\Sigma_2^{-1/2}$.
%\end{lemma}
To apply the above result to equation \eqref{eq_bias_2task}, we first approximate $X_1^{\top}X_1$ by $\Sigma_1$, and then use Lemma \ref{lem_cov_derivative} with $w = \Sigma_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)$.
Note that we can rescale $X_1$ by $W_1 / W_2$ and $\Sigma_1$ by $(W_1 / W_2)^2$, and the rest of the steps follows.
As a remark, the approximation of $X_1^{\top}X_1$ by $\Sigma_1$ incurs a generalization error that scales down with $\rho_1$, which is the $\delta / \sqrt{\rho_1}$ term in Theorem \ref{thm_main_informal} more precisely. The formal proof of Theorem \ref{thm_main_informal} using Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} will be presented in Section \ref{app_proof_main_thm}.
Finally, while the limiting bias $\Pi$ provides a precise connection with general covariate shift matrices $M$, it can be difficult to interpret.
In Section \ref{sec_special}, we show that $\Pi$ can be significantly simplified in an isotropic covariance setting.
%While the general form of these functions can be complex (as are previous generalization bounds for MTL), they admit interpretable forms for simplified settings.

@@

%which deals with the inverse of the sum of two random matrices, which
%any is can be viewed as a special case of Theorem \ref{thm_model_shift}.

%\begin{lemma}[Variance asymptotics]\label{lem_cov_shift}
	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	%Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
	%Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and
%	Let $\Sigma \in \real^{p \times p}$ be any fixed and deterministic matrix.
%	In the setting of Theorem \ref{thm_main_informal},
%	with probability $1-\oo(1)$ over the randomness of $X_1$ and $X_2$, we have that %for any constant $\e>0$,
	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
%	\begin{align}\label{lem_cov_shift_eq}
%		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma} = \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma} + o(\norm{\Sigma}), %\bigo{\|\| p^{-1/2+\epsilon}}
%	\end{align}
%where $(a_1,a_2)$ is the solution to equation \eqref{eq_a12extra} with $\lambda_i$ being the singular values of $M\equiv M(1): = \Sigma_1^{1/2}\Sigma_2^{-1/2}$.
%\end{lemma}

In the rest of section, we describe briefly the main ideas in the proof of Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}. In particular, we will give a simple (although not totally rigorous) derivation of the equations \eqref{eq_a12extra} and \eqref{eq_a34extra}.
The full proof of Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} will be given in Section \ref{appendix RMT}.
%\subsection{The Bias and Variance Asymptotics}

%\todo{merge} Finally, in this subsection, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.

\paragraph{Preliminaries.} Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $[(X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}]^{-1}$ for $X^{(1)}= Z^{(1)}\Sigma_1^{1/2}$ and $X^{(2)}= Z^{(2)}\Sigma_2^{1/2}$.
%\HZ{Suggestion: focus on the proof overview of Lemma \ref{lem_cov_shift} only. Use one paragraph in the end to say how to extend the idea to Lemma \ref{lem_cov_derivative}}
Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}[(X^{(1)})^\top (X^{(1)})+(X^{(2)})^\top (X^{(2)})]^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
%We divide the discussion into the following parts.
To study this expression, we shall use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. We now give some preliminaries about this method before discussing the proof strategy.

For any measure $\mu$ supported on $[0,\infty)$, it defines a Stieltjes transform through
$$m_\mu(z):= \int_0^\infty \frac{\dd\mu(x)}{x-z},\quad z\in \C\setminus [0,\infty).$$
Conversely, given a Stieltjes transform $m_{\mu}(z)$, we can recover the measure $\mu$ through the inverse formula
$$ \rho(x)=\frac{1}{\pi}\lim_{\eta \downarrow 0}\im m_{\mu}(x+\ii \eta),$$
where $\rho(x)$ is the density for $\mu$, i.e. $\dd \mu(x)=\rho(x)\dd x$. Hence the Stieltjes transform method reduces the study of a measure $\mu$ to the study of an analytic function $m_\mu(z)$. Let $\mu=p^{-1}\sum_{i=1}^p \delta_{\sigma_i}$ be the empirical spectral distribution of the matrix $Y_p:=(n_1+n_2)^{-1}(\Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V)$, where $\sigma_i$ denote the eigenvalues of $Y_p$, and the scaling $(n_1+n_2)^{-1}$ is chosen such that $\pi$ is supported on a bounded interval. Then with the above notations, the Stieltjes transform of the empirical spectral distribution $\mu$ is
$$ m_{\mu}=\frac{1}{p}\sum_{i=1}^p \frac{1}{\sigma_i - z}= p^{-1}\tr\left[ (Y_p-z\id_p)^{-1}\right].$$
The Stieltjes transform method in random matrix theory aims to show that $m_{\mu}$ satisfies a fixed-point equation asymptotically, which allows us to derive the asymptotic limit of $\mu$.
